DeepSphere: a graph-based spherical CNN

### 2019-09-25 main.tex
finish the proof

\paragraph{Theoretical convergence.}
We extend the work of Belkin and Nyiogi to a sufficiently regular, deterministic sampling scheme.
Given a sampling scheme $V = \{x_0, \dots, x_{n-1}\}$, define $\sigma_i$ to be the patch of the surface of the sphere corresponding to $x_i$, define $A_i$ to be its corresponding area and $d_i$ to be the radius of the smallest ball in $\R^3$ containing $\sigma_i$. Define $d^{(n)} := \max_{i=0, \dots, n}d_i$ and $A^{(n)}=\max_{i=0, \dots, n}A_i$.
\begin{theorem}
	For a sampling $V = \{x_i\in\S^2\}_{i=0}^{n-1}$ of the sphere that is equi-area and such that $d^{(n)} \leq \frac{C}{\sqrt{n}}$, for all $f: \S^2 \rightarrow \R$ Lipschitz with respect to the Euclidean distance in $\R^3$, for all $y\in\S^2$, there exists a sequence $t_n = n^\beta$ such that the rescaled Heat Kernel Graph Laplacian $\frac{|\S^2|}{4\pi t_n^2}\b {L}^t_n$ converges pointwise to the Laplace Beltrami operator on the sphere $\Delta{\S^2}$  for $n\to\infty$:
	$$ \lim_{n\to\infty} \frac{|\mathbb{S}^2|}{4\pi t_n^2} \b{L}_n^{t_n} f(x_i) =  \Delta_{\mathbb{S}^2}f(x_i)\quad \forall i=0, ..., n-1.$$
	\label{theo:pointwise convergence for a regular sampling}
\end{theorem}
This theorem implies immediately that $\b{L}$ converges pointwisely toward an operator ($\Delta_{\S^2}$) commuting with rotation and hence is a strong hint for equivariance.
Importantly, the proof of Theorem \ref{theo:pointwise convergence for a regular sampling} in Appendix \ref{sec: appendix: proof of theorem} inspires us in the construction of our graph Laplacian. It gives us a strong hint how the parameter $t$ needs to be adjusted with respect to the number of points $n$.

@@

The last proposition show that for a \emph{fixed} $t$, $L_n^tf(x)\rightarrow \linefrac{1}{4\pi^2} L^tf(x)$. To utilize Proposition \ref{prop:3} and complete the proof, we need to find a sequence of $t_n$ for which this holds as $t_n \rightarrow 0$. Furthermore this should hold with a faster decay than $\frac{1}{4\pi t_n^2}$.
\begin{prop}\label{prop:2}
	Given a sampling regular enough, i.e., for which we assume $A_i=A_j \ \forall i,j\text{ and }d^{(n)}\leq \frac{C}{\sqrt{n}}$, a Lipschitz function $f$ and a point $y\in\S^2$ there exists a sequence $t_n = n^\beta, \beta<0$ such that
\begin{equation*}
    \forall f \text{ Lipschitz, } \forall x\in\S^2 \quad \left|\frac{1}{4\pi t_n^2}\left(L_n^{t_n}f(x) -  L^{t_n}f(x)\right)\right|\xrightarrow{n\to \infty}0.
\end{equation*}
\end{prop}
\begin{proof}
% [Proof of Proposition \ref{prop:2}]
To ease the notation, we define
\begin{align}
	K^t(x,y) &:=  e^{-\frac{\|x-y\|^2}{4t}}\\
	\phi^t(x;y) &:= e^{-\frac{\|x-y\|^2}{4t}}\left(f(y)-f(x)\right).
\end{align}
We start with the following inequality
\begin{align*}
	\|L_n^tf-L^tf\|_\infty &= \max _{y\in \S^2} \left|L_n^tf(y)-L^tf(y)\right|\\
	&= \max _{y\in \S^2} \left| \frac{1}{n} \sum_{i=1}^n \phi^t(x_i; y)- \int_{\S^2} \phi^t(x;y)d\mu(x) \right|\\
	&\leq \max _{y\in \S^2}  \sum_{i=1}^n   \left| \frac{1}{n}  \phi^t(x_i; y)- \int_{\sigma_i} \phi^t(x;y)d\mu(x) \right|\\
	&\leq  d^{(n)} \max _{y\in \S^2} C_{\phi^t_y} ,\\
\end{align*}
where $C_{\phi^t_y}$ is the Lipschitz constant of $x \rightarrow \phi^t(x, y)$ and the last inequality follows from Proposition \ref{prop:1}. 
Using the assumption $d^{(n)}\leq \frac{C}{\sqrt{n}}$ we find
\begin{equation*}
\|L_n^tf-L^tf\|_\infty  \leq  \frac{C}{\sqrt{n}} \max _{y\in \S^2}  C_{\phi^t_y}  
\end{equation*}
We now find the explicit dependence between $t$ and $C_{\phi^t_y}$
\begin{align*}
	C_{\phi^t_y} &= \|\partial_x\phi^t(\cdot;y)|\_\infty\\&
	= \|\partial_x\left(K^t(\cdot;y)f\right)\|_\infty\\&
	= \|\partial_x K^t(\cdot;y)f + K^t(\cdot;y)\partial_x f||_\infty\\&
	\leq \|\partial_x K^t(\cdot;y)f\|_\infty + \|K^t(\cdot;y)\partial_x f\|_\infty\\&
	\leq  \|\partial_x K^t(\cdot;y)\|_\infty\|f\|_\infty + \|K^t(\cdot;y)\|_\infty\|\partial_x f\|_\infty\\&
	= \|\partial_x K^t(\cdot;y)\|_\infty\|f\|_\infty + \|\partial_x f\|_\infty\\&
	= C_{K^t_y} \|f\|_\infty + \|\partial_xf\|_\infty\\&
	= C_{K^t_y} \|f\|_\infty + C_f
\end{align*}
where $C_{K^t_y}$ is the Lipschitz constant of $x\rightarrow K^t(x;y)$. We can observe that such constant does not depend on $y$:

$\mathcal L_{K^t_y} = \norm{\partial_x e^{-\frac{x^2}{4t}}}_\infty = \norm{\frac{x}{2t}e^{-\frac{x^2}{4t}}}_\infty = \left. \frac{x}{2t}e^{-\frac{x^2}{4t}}\right|_{x=\sqrt{2t}}=(2et)^{-\frac{1}{2}}\propto t ^ {-\frac{1}{2}}$

So we can continue
$$\begin{aligned}
	\max _{y\in \S^2} \left[  \mathcal L_{\phi^t_y} \frac{C}{\sqrt{n}} \right]
	&\leq  \frac{C}{\sqrt{n}} \left( (2et)^{-\frac{1}{2}} \norm{f}_\infty + \mathcal L_f \right)\\
	&\leq \frac{C\norm{f}_\infty}{\sqrt{n}(2et)^{1/2}} +   \frac{C}{\sqrt{n}}\mathcal L_f\\
\end{aligned}$$
So we have that, rescaling by a factor $\frac{1}{4\pi t^2}$
\begin{align*}
	\norm{\frac{1}{4\pi t^2}\left(L_n^tf-L^tf\right)}_\infty&\leq \frac{1}{4\pi t^2}\norm{\left(L_n^tf-L^tf\right)}_\infty \\
	&\leq \frac{C}{4\pi}\left[\frac{\norm{f}_\infty}{\sqrt{2e}}\frac{1}{\sqrt{n}t^{5/2}} + \frac{\mathcal L_f}{\sqrt{n}t^2}\right]
\end{align*}

we want $\begin{cases}
t \rightarrow 0\\
n \rightarrow \infty\\
\sqrt{n}t^{5/2} \rightarrow \infty\\
\sqrt{n}t^2 \rightarrow \infty
\end{cases}$ in order for $ \frac{C}{4\pi}\left[\frac{\norm{f}_\infty}{\sqrt{2e}}\frac{1}{\sqrt{n}t^{5/2}} + \frac{\mathcal L_f}{\sqrt{n}t^2}\right] \xrightarrow[t\to 0 ]{n\to\infty}0$

This is true if $\begin{cases}
t(n) = n^\beta, &\beta\in(-\frac{1}{5}, 0) \\
t(n) = n^\beta, &\beta\in(-\frac{1}{4}, 0)
\end{cases} \implies t(n) = n^\beta, \quad \beta\in(-\frac{1}{5}, 0)$

Indeed

$\sqrt{n}t^{5/2}=n^{5/2\beta+1/2}\xrightarrow{n \to \infty} \infty$ since $\frac{5}{2}\beta+1/2>0 \iff \beta>-\frac{1}{5}$

$\sqrt{n}t^2=n^{2\beta+1/2}\xrightarrow {N \to \infty} \infty$ since $2\beta+1/2>0 \iff \beta>-\frac{1}{4}$

So, for $t=n^\beta$ with $\beta\in(-\frac{1}{5}, 0)$ we have that

$$\begin{cases}
(t_n)\xrightarrow{n\to\infty}0\\
\norm{\frac{1}{4\pi t_n^2}L_n^{t_n}f-\frac{1}{4\pi t_n^2}L^{t_n}f}_\infty  \xrightarrow{n\to\infty}0
\end{cases}$$

@@

Theorem  \ref{theo:pointwise convergence for a regular sampling}, is then an immediate consequence of Proposition \ref{prop:2} and Proposition \ref{prop:3}.
\begin{proof}[Proof of Theorem \ref{theo:pointwise convergence for a regular sampling}]
	Thanks to Proposition \ref{prop:2} and Proposition \ref{prop:3}	we conclude that $\forall y\in\S^2 $
	$$\lim_{n\to\infty}\frac{1}{4\pi t_n^2} L_n^{t_n}f(y) =  \lim_{n\to\infty}\frac{1}{4\pi t_n^2} L^{t_n}f(y) = \frac{1}{|\S^2|}\triangle_{\S^2}f(y) $$
\end{proof}

>>>>

\paragraph{Theoretical convergence.}
We extend the work of Belkin and Nyiogi to a sufficiently regular, deterministic sampling scheme.
Given a sampling scheme $V = \{x_0, \dots, x_{n-1}\}$, define $\sigma_i$ to be the patch of the surface of the sphere corresponding to $x_i$, define $A_i$ to be its corresponding area and $d_i$ to be the radius of the smallest ball in $\R^3$ containing $\sigma_i$. Define $d^{(n)} := \max_{i=0, \dots, n}d_i$ and $A^{(n)}=\max_{i=0, \dots, n}A_i$.
\begin{theorem}
	For a sampling $V = \{x_i\in\S^2\}_{i=0}^{n-1}$ of the sphere that is equi-area and such that $d^{(n)} \leq \frac{C}{\sqrt{n}}$, for all $f: \S^2 \rightarrow \R$ Lipschitz with respect to the Euclidean distance in $\R^3$, for all $y\in\S^2$, there exists a sequence $t_n = n^\beta$ such that the rescaled Heat Kernel Graph Laplacian $\frac{|\S^2|}{4\pi t_n^2}\b {L}^t_n$ converges pointwise to the Laplace Beltrami operator on the sphere $\Delta{\S^2}$  for $n\to\infty$:
	\begin{equation*}
	 \lim_{n\to\infty} \frac{|\mathbb{S}^2|}{4\pi t_n^2} \b{L}_n^{t_n} f(x_i) =  \Delta_{\mathbb{S}^2}f(x_i)\quad \forall i=0, ..., n-1.
	\end{equation*}
	\label{theo:pointwise convergence for a regular sampling}
\end{theorem}
This theorem implies immediately that $\b{L}$ converges pointwisely toward an operator ($\Delta_{\S^2}$) commuting with rotation and hence is a strong hint for equivariance.
Importantly, the proof of Theorem \ref{theo:pointwise convergence for a regular sampling} in Appendix \ref{sec: appendix: proof of theorem} inspires us in the construction of our graph Laplacian. It gives us a strong hint how the parameter $t$ needs to be adjusted with respect to the number of points $n$.

@@

The last proposition show that for a \emph{fixed} $t$, $L_n^tf(x)\rightarrow \linefrac{1}{4\pi^2} L^tf(x)$. To utilize Proposition \ref{prop:3} and complete the proof, we need to find a sequence of $t_n$ for which this holds as $t_n \rightarrow 0$. Furthermore this should hold with a faster decay than $\frac{1}{4\pi t_n^2}$.
\begin{prop}\label{prop:2}
	Given a sampling regular enough, i.e., for which we assume $A_i=A_j \ \forall i,j\text{ and }d^{(n)}\leq \frac{C}{\sqrt{n}}$, a Lipschitz function $f$ and a point $y\in\S^2$ there exists a sequence $t_n = n^\beta, \beta<0$ such that
\begin{equation*}
    \forall f \text{ Lipschitz, } \forall x\in\S^2 \quad \left|\frac{1}{4\pi t_n^2}\left(L_n^{t_n}f(x) -  L^{t_n}f(x)\right)\right|\xrightarrow{n\to \infty}0.
\end{equation*}
\end{prop}
\begin{proof}
% [Proof of Proposition \ref{prop:2}]
To ease the notation, we define
\begin{align}
	K^t(x,y) &:=  e^{-\frac{\|x-y\|^2}{4t}}\\
	\phi^t(x;y) &:= e^{-\frac{\|x-y\|^2}{4t}}\left(f(y)-f(x)\right).
\end{align}
We start with the following inequality
\begin{align}
	\|L_n^tf-L^tf\|_\infty &= \max _{y\in \S^2} \left|L_n^tf(y)-L^tf(y)\right| \nonumber\\
	&= \max _{y\in \S^2} \left| \frac{1}{n} \sum_{i=1}^n \phi^t(x_i; y)- \int_{\S^2} \phi^t(x;y)d\mu(x) \right| \nonumber\\
	&\leq \max _{y\in \S^2}  \sum_{i=1}^n   \left| \frac{1}{n}  \phi^t(x_i; y)- \int_{\sigma_i} \phi^t(x;y)d\mu(x) \right| \nonumber\\
	&\leq  d^{(n)} \max _{y\in \S^2} C_{\phi^t_y} , \label{eq:prop3-base-ineq}
\end{align}
where $C_{\phi^t_y}$ is the Lipschitz constant of $x \rightarrow \phi^t(x, y)$ and the last inequality follows from Proposition \ref{prop:1}. 
Using the assumption $d^{(n)}\leq \frac{C}{\sqrt{n}}$ we find
\begin{equation*}
\|L_n^tf-L^tf\|_\infty  \leq  \frac{C}{\sqrt{n}} \max _{y\in \S^2}  C_{\phi^t_y}  
\end{equation*}
We now find the explicit dependence between $t$ and $C_{\phi^t_y}$
\begin{align*}
	C_{\phi^t_y} &= \|\partial_x\phi^t(\cdot;y)|\_\infty\\&
	= \|\partial_x\left(K^t(\cdot;y)f\right)\|_\infty\\&
	= \|\partial_x K^t(\cdot;y)f + K^t(\cdot;y)\partial_x f||_\infty\\&
	\leq \|\partial_x K^t(\cdot;y)f\|_\infty + \|K^t(\cdot;y)\partial_x f\|_\infty\\&
	\leq  \|\partial_x K^t(\cdot;y)\|_\infty\|f\|_\infty + \|K^t(\cdot;y)\|_\infty\|\partial_x f\|_\infty\\&
	= \|\partial_x K^t(\cdot;y)\|_\infty\|f\|_\infty + \|\partial_x f\|_\infty\\&
	= C_{K^t_y} \|f\|_\infty + \|\partial_xf\|_\infty\\&
	= C_{K^t_y} \|f\|_\infty + C_f
\end{align*}
where $C_{K^t_y}$ is the Lipschitz constant of the function $x\rightarrow K^t(x;y)$. We note that this constant does not depend on $y$:
\begin{equation*}
C_{K^t_y} = \norm{\partial_x e^{-\frac{x^2}{4t}}}_\infty = \norm{\frac{x}{2t}e^{-\frac{x^2}{4t}}}_\infty = \left. \frac{x}{2t}e^{-\frac{x^2}{4t}}\right|_{x=\sqrt{2t}}=(2et)^{-\frac{1}{2}}\propto t ^ {-\frac{1}{2}}.
\end{equation*}
Hence we have
\begin{align*}
	\frac{C}{\sqrt{n}}  \max _{y\in \S^2} C_{\phi^t_y} 
	&\leq  \frac{C}{\sqrt{n}} \left( (2et)^{-\frac{1}{2}} \norm{f}_\infty + C_f \right)\\
	&\leq \frac{C \norm{f}_\infty}{\sqrt{n}(2et)^{1/2}} +   \frac{C}{\sqrt{n}} C_f.
\end{align*}
Inculding this result in \eqref{eq:prop3-base-ineq} and rescaling by $\linefrac{1}{4\pi t^2}$, we obtain
\begin{align*}
	\norm{\frac{1}{4\pi t^2}\left(L_n^tf-L^tf\right)}_\infty&\leq \frac{1}{4\pi t^2}\norm{\left(L_n^tf-L^tf\right)}_\infty \\
	&\leq \frac{C}{4\pi}\left[\frac{\norm{f}_\infty}{\sqrt{2e}}\frac{1}{\sqrt{n}t^{5/2}} + \frac{C_f}{\sqrt{n}t^2}\right].
\end{align*}
In order for $ \frac{C}{4\pi}\left[\frac{\norm{f}_\infty}{\sqrt{2e}}\frac{1}{\sqrt{n}t^{5/2}} + \frac{C_f}{\sqrt{n}t^2}\right] \xrightarrow[t\to 0 ]{n\to\infty}0$, 
we need $\begin{cases}
% t \rightarrow 0\\
% n \rightarrow \infty\\
\sqrt{n}t^{5/2} \rightarrow \infty\\
\sqrt{n}t^2 \rightarrow \infty
\end{cases}$ \\
It happens if $\begin{cases}
t(n) = n^\beta, &\beta\in(-\frac{1}{5}, 0) \\
t(n) = n^\beta, &\beta\in(-\frac{1}{4}, 0)
\end{cases} \implies t(n) = n^\beta, \quad \beta\in(-\frac{1}{5}, 0)$.\\
Indeed, we have\\
$\sqrt{n}t^{5/2}=n^{5/2\beta+1/2}\xrightarrow{n \to \infty} \infty$ since $\frac{5}{2}\beta+1/2>0 \iff \beta>-\frac{1}{5}$\\
and $\sqrt{n}t^2=n^{2\beta+1/2}\xrightarrow {N \to \infty} \infty$ since $2\beta+1/2>0 \iff \beta>-\frac{1}{4}$.\\
As a result, for $t=n^\beta$ with $\beta\in(-\frac{1}{5}, 0)$ we have
$\begin{cases}
(t_n)\xrightarrow{n\to\infty}0\\
\norm{\frac{1}{4\pi t_n^2}L_n^{t_n}f-\frac{1}{4\pi t_n^2}L^{t_n}f}_\infty  \xrightarrow{n\to\infty}0,
\end{cases}$\\
which concludes the proof.

@@

Theorem  \ref{theo:pointwise convergence for a regular sampling}, is then an immediate consequence of Proposition \ref{prop:2} and Proposition \ref{prop:3}.
\begin{proof}[Proof of Theorem \ref{theo:pointwise convergence for a regular sampling}]
	Thanks to Proposition \ref{prop:2} and Proposition \ref{prop:3}	we conclude that $\forall y\in\S^2 $
	\begin{equation*}
	\lim_{n\to\infty}\frac{1}{4\pi t_n^2} L_n^{t_n}f(y) =  \lim_{n\to\infty}\frac{1}{4\pi t_n^2} L^{t_n}f(y) = \frac{1}{|\S^2|}\triangle_{\S^2}f(y)
	\end{equation*}
	Eventually selecting $y$ as a point of the sampling, we obtain
	\begin{equation*}
	 \lim_{n\to\infty} \frac{|\mathbb{S}^2|}{4\pi t_n^2} \b{L}_n^{t_n} f(x_i) =  \Delta_{\mathbb{S}^2}f(x_i)\quad \forall i=0, ..., n-1.
	\end{equation*} 
\end{proof}
